Syntax and Semantics of Cedille

### 2020-12-30 docs/semantics/paper.tex
docs: semantics: finish validity
- make Lemma 15 mutual with a similar proof for term variables in context
- update existing cases with new rules

  \item If \(\Gamma_1,X:\kappa_1,\Gamma_2 \vdash T \tpsynth \kappa_2\) and
    \(\Gamma_1 \vdash \kappa_1'\) then \(\Gamma_1,X:\kappa_1',\Gamma_2 \vdash T
    \tpsynth \kappa_2\) for some \(\kappa_2' \cong \kappa_2\)

  \item If \(\Gamma_1,X:\kappa_1,\Gamma_2 \vdash t \Leftrightarrow T\) and
    \(\Gamma_1 \vdash \kappa_1'\) then \(\Gamma_1,X:\kappa_1',\Gamma_2 \vdash t
    \Leftrightarrow T\).
  \end{itemize}
\end{lemma}

\begin{proof}
  By induction on the assumed derivation.
  We show a few interesting cases.

  \startcase{.2cm}
  \[
    \infer{\Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash X \tpsynth \kappa}{(X : \kappa) \in \Gamma_1,X_1:\kappa_1,\Gamma_2}
  \]
  We have two subcases to consider.
  If \(X = X_1\), then by the rule we obtain \(\Gamma_1,X_1:\kappa_1',\Gamma_2
  \vdash X_1 \tpsynth \kappa_1'\), as desired.
  Otherwise, by the rule we obtain \(\Gamma_1,X_1:\kappa_1',\Gamma_2
  \vdash X \tpsynth \kappa\).

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash \abs{\lambda}{X}{\kappa}{T'} \tpsynth
      \abs{\Pi}{X}{\kappa}{\kappa'}
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash \kappa
      & \Gamma_1,X_1:\kappa_1,\Gamma_2,X:\kappa \vdash  T' \tpsynth \kappa'
    }
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash
  \kappa\).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2,X:\kappa
  \vdash T' \tpsynth \kappa''\) for some \(\kappa'' \cong \kappa'\).
  By the rule, \(\Gamma_1,X_1\kappa_1',\Gamma_2 \vdash
  \abs{\lambda}{X}{\kappa}{T'} \tpsynth \abs{\Pi}{X}{\kappa}{\kappa''}\), which
  is convertible with the original kind synthesized for this type.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T \cdot T' \tpsynth
      [T'/X]\kappa
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T \tpsynth
      \abs{\Pi}{X}{\kappa'}{\kappa}
      & \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T' \tpsynth \kappa''
      & \kappa' \cong \kappa''}
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T
  \tpsynth \abs{\Pi}{X}{\kappa_2'}{\kappa_2}\) for some \(\kappa_2' \cong
  \kappa'\), \(\kappa_2 \cong \kappa\).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T'
  \tpsynth \kappa_2''\) for some \(\kappa_2'' \cong \kappa''\).
  By transitivity, \(\kappa_2' \cong \kappa_2''\).
  By the rule, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash T \cdot T' \tpsynth
  [T'/X]\kappa_2\), with this kind congruent to \([T'/X]\kappa\).

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t \cdot T' \tpsynth [T'/X]T
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2\vdash t \tpsynth
      \abs{\forall}{X}{\kappa}{T}
      & \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T' \tpsynth \kappa'
      & \kappa'\cong\kappa
    }
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t
  \tpsynth \abs{\forall}{X}{\kappa}{T}\).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash T'
  \tpsynth \kappa''\) for some \(\kappa'' \cong \kappa'\).
  With the third premise and transitivity of congruence, we have \(\kappa''
  \cong \kappa\).
  We apply the rule and conclude.
\end{proof}

\begin{corollary}
  \label{cor:ctxt-conv-class}
  If \(\vdash \Gamma_1,X:\kappa_1,\Gamma_1\) and \(\Gamma_1 \vdash \kappa_1'\)
  with \(\kappa_1' \cong \kappa\) then \(\vdash \Gamma_1,X:\kappa_1',\Gamma_2\).
\end{corollary}

\begin{lemma}
  \label{lem:subst-class}
  Below, each statement seperately universally quantifies over meta-variables,
  and it is assumed that typing contexts occuring in assumed derivations are
  well-formed.
  \begin{enumerate}
  \item \textbf{Kinds:}
    \begin{itemize}
    \item If \(\Gamma_1,x:T,\Gamma_2 \vdash \kappa\) and \(\Gamma_1 \vdash t
      \tpcheck T\) then \(\Gamma_1,[t/x]\Gamma_2 \vdash [t/x] \kappa\)

>>>>

  \item If \(\Gamma_1,X:\kappa_1,\Gamma_2 \vdash T \tpsynth \kappa_2\) and
    \(\Gamma_1 \vdash \kappa_1'\) then \(\Gamma_1,X:\kappa_1',\Gamma_2 \vdash T
    \tpsynth \kappa'_2\) for some \(\kappa_2' \cong \kappa_2\)

  \item If \(T_3 \cong T_4\) and \(\Gamma_1,X:\kappa_1,\Gamma_2 \vdash t
    \tpcheck T_3\) then \(\Gamma_1,X:\kappa_1',\Gamma_2 \vdash t \tpcheck T_4\)

  \item If \(\Gamma_1,X:\kappa_1,\Gamma_2 \vdash t \tpsynth T\) and
    \(\Gamma_1 \vdash \kappa_1'\) then \(\Gamma_1,X:\kappa_1',\Gamma_2 \vdash t
    \tpsynth T\).
  \end{itemize}
  Furthermore, the resulting typing derivations derivations have depths no
  larger than the assumed ones.
\end{lemma}

\begin{proof}
  By mutual induction on the assumed derivation, and mutually with
  Lemma~\ref{lem:ctxt-conv-class2}.
  The measure, which omits the size of derivations of the convertibility
  relation, is used to ensure that each mutually inductive call is well-founded.
  We show the interesting cases.

  % kinding
  \startcase{.2cm}
  \[
    \infer{\Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash X \tpsynth \kappa}{(X : \kappa) \in \Gamma_1,X_1:\kappa_1,\Gamma_2}
  \]
  We have two subcases to consider.
  If \(X = X_1\), then by the rule we obtain \(\Gamma_1,X_1:\kappa_1',\Gamma_2
  \vdash X_1 \tpsynth \kappa_1'\), as desired.
  Otherwise, by the rule we obtain \(\Gamma_1,X_1:\kappa_1',\Gamma_2
  \vdash X \tpsynth \kappa\).
  In both cases, the resulting derivation has the same depth (i.e., 1) as we
  started with.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash\abs{\lambda}{x}{T}{T'} \tpsynth
      \abs{\Pi}{x}{T}{\kappa}
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T \tpsynth \star
      & \Gamma_1,X_1:\kappa_1,\Gamma_2 ,x:T\vdash T'\tpsynth\kappa
    }
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash T
  \tpsynth \star\) at no greater depth (note \(\star\) is convertible only with
  itself).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2,x:T \vdash
  T' \tpsynth \kappa'\) (at no greater depth) for some \(\kappa' \cong \kappa\).
  By the rule, we have \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash
  \abs{\lambda}{x}{T}{T'} \tpsynth \abs{\Pi}{x}{T}{\kappa'}\), which is
  congruent with the original type in the conclusion.
  We see the the measure is preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash \abs{\lambda}{X}{\kappa}{T'} \tpsynth
      \abs{\Pi}{X}{\kappa}{\kappa'}
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash \kappa
      & \Gamma_1,X_1:\kappa_1,\Gamma_2,X:\kappa \vdash  T' \tpsynth \kappa'
    }
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash
  \kappa\).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2,X:\kappa
  \vdash T' \tpsynth \kappa''\) for some \(\kappa'' \cong \kappa'\).
  By the rule, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash
  \abs{\lambda}{X}{\kappa}{T'} \tpsynth \abs{\Pi}{X}{\kappa}{\kappa''}\), which
  is convertible with the original kind synthesized for this type.
  Since the depths of the premises for the resulting derivation are the same as
  the depths for the premises of the assumed one, the measure is preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T\ t \tpsynth [t/x]^{T'}\kappa
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T \tpsynth \abs{\Pi}{x}{T'}{\kappa}
      & \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t \tpcheck T'
    }
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash T
  \tpsynth \abs{\Pi}{x}{T''}{\kappa''}\) for some \(T'' \cong T'\) and
  \(\kappa'' \cong \kappa\).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t
  \tpcheck T''\).
  By the rule, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash T\ t \tpsynth
  [t/x]^{T''}\kappa''\), convertible with the given type \([t/x]^{T'}\kappa\).
  We see that the measure is preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T \cdot T' \tpsynth
      [T'/X]\kappa
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T \tpsynth
      \abs{\Pi}{X}{\kappa'}{\kappa}
      & \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T' \tpsynth \kappa''
      & \kappa' \cong \kappa''}
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T
  \tpsynth \abs{\Pi}{X}{\kappa_2'}{\kappa_2}\) for some \(\kappa_2' \cong
  \kappa'\), \(\kappa_2 \cong \kappa\).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T'
  \tpsynth \kappa_2''\) for some \(\kappa_2'' \cong \kappa''\).
  By transitivity and the third premise, \(\kappa_2' \cong \kappa_2''\).
  By the rule, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash T \cdot T' \tpsynth
  [T'/X]\kappa_2\), with this kind convertible to \([T'/X]\kappa\) by
  congruence.
  We see the measure is preserved.

  % checking
  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t\tpcheck T_3
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t\tpsynth T'
      & T' \cong T_3
    }
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t
  \tpsynth T'\).
  By assumption and transitivity of congruence, \(T' \cong T_4\) (where \(T_4\)
  is the type given to us in the proof).
  By the rule, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t \tpcheck T_4\), and we
  see the measure is preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash \absu{\lambda}{x}{t} \tpcheck T_3
    }{
      T_3 \leadsto_\beta^* \abs{\Pi}{x}{T_1}{T_2}
      & \Gamma_1,X_1:\kappa_1,\Gamma_2,x:T_1\vdash t\tpcheck T_2
    }
  \]
  By the first premise, \(T_3 \cong \abs{\Pi}{x}{T_1}{T_2}\).
  By assumption and transitivity of the convertibility relation, \(T_4 \cong
  \abs{\Pi}{x}{T_1}{T_2}\).
  This means \(T_4 \leadsto_\beta^* \abs{\Pi}{x}{T_1'}{T_2'}\) for some \(T_1'
  \cong T_1\) and \(T_2' \cong T_2\).
  By mutual induction with Lemma~\ref{lem:ctxt-conv-class2} on the second
  premise, we have a derivation of \(\Gamma_1,X_1:\kappa_1,\Gamma_2,x:T_1'
  \vdash t \tpcheck T_2'\) that is no deeper than that premise.
  So, we are entitled to use the IH on this new derivation to obtain
  \(\Gamma_1,X_1:\kappa_1',\Gamma_2,x:T_1' \vdash t \tpcheck T_2'\), which is
  also no deeper.
  By the rule, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash \absu{\lambda}{x}{t}
  \tpcheck T_4\), and the measure is preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2\vdash \absu{\Lambda}{X}{t} \tpcheck T_3
    }{
      T_3 \leadsto_\beta^* \abs{\forall}{X}{\kappa}{T}
      & \Gamma_1,X_1:\kappa_1,\Gamma_2,X:\kappa\vdash t \tpcheck T
    }
  \]
  By the first premise, \(T_3 \cong \abs{\forall}{X}{\kappa}{T}\), so by
  assumption and transitivity of convertibility, \(T_4 \cong
  \abs{\forall}{X}{\kappa}{T}\).
  This means \(T_4 \leadsto_\beta^* \abs{\forall}{X}{\kappa'}{T'}\) for some
  \(\kappa' \cong \kappa\) and \(T' \cong T\).
  We apply the IH once on the second premise to obtain a derivation of
  \(\Gamma_1,X_1:\kappa_1,\Gamma_2,X:\kappa' \vdash t \tpcheck T'\), and then
  again to obtain \(\Gamma_1,X_1:\kappa_1',\Gamma_2,X:\kappa' \vdash t \tpcheck
  T'\), noting this derivation's depth is no greater than the second premise.
  By the rule, \(\Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash \absu{\Lambda}{X}{t}
  \tpcheck T_4\), and the measure is preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash \beta\{t'\} \tpcheck T_3
    }{
      T_3 \leadsto_\beta^* \{t_1 \simeq t_2\}
      & \textit{FV}(t')\subseteq \textit{dom}(\Gamma_1,X_1:\kappa_1,\Gamma_2)
      & |t_1| =_{\beta\eta} |t_2|
    }
  \]
  From the first premise, assumption, and transitivity of convertibility, \(T_4
  \cong \{t_1 \simeq t_2\}\).
  This means that \(T_4 \leadsto_\beta^* \{t_3 \simeq t_4\}\) for some
  \(t_3,t_4\) such that \(|t_3| =_{\beta\eta} |t_1|\) and \(|t_4| =_{\beta\eta}
  |t_2|\).
  From this, the third premise, and transitivity of convertibility, we obtain
  \(|t_3| =_{\beta\eta} |t_4|\).
  We also see from the second premise that \(\textit{FV}(t') \subseteq
  \textit{dom}(\Gamma_1,X_1:\kappa_1',\Gamma_2)\).
  So, we use the rule to conclude that \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash
  \beta\{t'\} \tpcheck T_4\), and the depth (1) is preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash [ t_1 , t_2 ] \tpcheck T_3
    }{
      \begin{array}{cc}
        T_3 \leadsto_\beta^* \abs{\iota}{x}{T_1}{T_1}
        & \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t_1 \tpcheck T_1
        \\ \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t_2 \tpcheck [t/x]^{T_1}\ T_2
        & |t_1| =_{\beta\eta} |t_2|
      \end{array}
    }
  \]
  From the first premise, assumption, and transitivity of the convertibility
  relation, \(T_4 \cong \abs{\iota}{x}{T_1}{T_2}\).
  This means that \(T_4 \leadsto_\beta^* \abs{\iota}{x}{T_1'}{T_2'}\) for some
  \(T_1' \cong T_1\) and \(T_2' \cong T_2\).
  From this last congruence, we have \([t_1/x]^{T_1'}T_2' \cong [t_1/x]^{T_1} T_2\).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t_1
  \tpcheck T_1'\).
  By the IH on the first premise, second row, we have
  \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t_2 \tpcheck [t/x]T_2'\).
  We use the rule to conclude, noting that the measure is preserved.

  % synthesis
  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t\ t' \tpsynth [t'/x]^{T'}T
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t \tpsynthleads \abs{\Pi}{x}{T'}{T}
      & \Gamma_1,X_1:\kappa_1,\Gamma_2\vdash t' \tpcheck T'
    }
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t
  \tpsynthleads \abs{\Pi}{x}{T'}{T}\).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t'
  \tpcheck T'\).
  By the rule, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t\ t' \tpsynth [t'/x]T\),
  and the measure is preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t \cdot T' \tpsynth [T'/X]T
    }{
      \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t \tpsynthleads
      \abs{\forall}{X}{\kappa}{T}
      & \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash T' \tpsynth \kappa'
      & \kappa'\cong\kappa
    }
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t
  \tpsynthleads \abs{\forall}{X}{\kappa}{T}\).
  By the IH on the second premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash T'
  \tpsynth \kappa''\) for some \(\kappa'' \cong \kappa'\), from which we obtain
  \(\kappa'' \cong \kappa\).
  By the rule, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t \cdot T' \tpsynth
  [T'/X]T\), and the measure is preserved.

  \startcase{.2cm}
  \[
    \infer{\Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash \rho\ t'\ @x\langle t_2 \rangle.T\ \mhyph\ t \tpsynth [t_2/x]\ T}
    {
      \begin{array}{ccc}
        \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t' \tpsynthleads \{t_1 \simeq t_2'\}
        & \textit{FV}(t_2) \subseteq \textit{dom}(\Gamma_1,X_1:\kappa_1,\Gamma_2)
        & |t_2'| =_{\beta\eta} |t_2|
        \\ \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash [t_2/x]\ T \tpsynth \star
        & \Gamma_1,X_1:\kappa_1,\Gamma_2 \vdash t \tpsynth T'
        & T' \cong [t_1/x]T
      \end{array}
    }
  \]
  By the IH on the first premise, \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash t'
  \tpsynthleads \{t_1 \simeq t_2'\}\).
  By the IH on the second premise of the second row,
  \(\Gamma_1,X_1:\kappa_1',\Gamma_2 \vdash [t_2/x]\ T \tpsynth \star\)
  (\(\star\) is only convertible with itself).
  We apply the IH to the second premise of the second row and use the rule to conclude
  (\(\textit{dom}(\Gamma_1,X_1:\kappa_1,\Gamma_2) =
  \textit{dom}(\Gamma_1,X_1:\kappa_1',\Gamma_2)\)).
\end{proof}

\begin{lemma}
  \label{lem:ctxt-conv-class2}
  Let \(T_1,T_2\) be types such that \(T_1 \cong T_2\).
  \begin{itemize}
  \item If \(\Gamma_1,x:T_1,\Gamma_2 \vdash \kappa\) then
    \(\Gamma_1,x:T_2,\Gamma_2 \vdash \kappa\)

  \item If \(\Gamma_1,x:T_1,\Gamma_2 \vdash T \tpsynth \kappa_1\) then
    \(\Gamma_1,x:T_2,\Gamma_2 \vdash T \tpsynth \kappa_1\)

  \item If \(T_3 \cong T_4\) and \(\Gamma_1,x:T_1,\Gamma_2 \vdash t \tpcheck T_3\) then
    \(\Gamma_1,x:T_2,\Gamma_2 \vdash t \tpcheck T_4\)

  \item If \(\Gamma_1,x:T_1,\Gamma_2 \vdash t \tpsynth T\) then
    \(\Gamma_1,x:T_2,\Gamma_2 \vdash t \tpsynth T'\) for some \(T' \cong T\).

  \end{itemize}
  Furthermore, the resulting typing derivations derivations have depths no
  larger than the assumed ones.
\end{lemma}
\begin{proof}
  By mutual induction on the assumed derivation (specifically, its depth), and
  mutually with Lemma~\ref{lem:ctxt-conv-class}.
  The measure, which omits the size of derivations of the convertibility
  relation, is used to ensure that each mutually inductive call is well-founded.
  We show the interesting cases.

  % kinding types
  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash \abs{\lambda}{x}{T}{T'} \tpsynth \abs{\Pi}{x}{T}{\kappa}
    }{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash T \tpsynth \star
      & \Gamma_1,x_1:T_1,\Gamma_2,x:T \vdash T' \tpsynth \kappa
    }
  \]
  By the IH on the first premise, \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash T \tpsynth
  \star\).
  By the IH on the second premise, \(\Gamma_1,x_1:T_2,\Gamma_2,x:T \vdash T'
  \tpsynth \kappa\).
  We also have that the depths of the resulting typing derivations are no greater than
  the those used to derive the premises, so we conclude using the rule.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,x_1:T_1,\Gamma_2\vdash\abs{\lambda}{X}{\kappa}{T'} \tpsynth
      \abs{\Pi}{X}{\kappa}{\kappa'}
    }{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash \kappa & \Gamma_1,x_1:T_1,\Gamma_2,X:\kappa\vdash T'\tpsynth\kappa'
    }
  \]
  By the IH on the first premise, \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash T \tpsynth
  \kappa\).
  By the IH on the second premise, \(\Gamma_1,x_1:T_2,\Gamma_2,X:\kappa \vdash
  T' \tpsynth \kappa'\).
  We also have that the depths of the resulting derivations are no greater than
  the those used to derive the premises, so we conclude using the rule.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,x_1:T_1,\Gamma_2\vdash T\ t \tpsynth [t/x]^{T'}\kappa
    }{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash T \tpsynth \abs{\Pi}{x}{T'}{\kappa}
      & \Gamma_1,x_1:T_1,\Gamma_2 \vdash t \tpcheck T'
    }
  \]
  By the IH on the first premise, \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash T \tpsynth
  \abs{\Pi}{x}{T'}{\kappa}\).
  By the IH on the second premise, \(\Gamma_2,x_1:T_2,\Gamma_2 \vdash t \tpcheck
  T'\).
  We also have that the depths of the resulting derivations are no greater than
  the those used to derive the premises, so we conclude using the rule.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash T \cdot T' \tpsynth [T'/X]\kappa
    }{
      \Gamma_1,x_1:T_1,\Gamma_2\vdash T \tpsynth \abs{\Pi}{X}{\kappa'}{\kappa}
      & \Gamma_1,x_1:T_1,\Gamma_2 \vdash T_2 \tpsynth \kappa'' & \kappa'' \cong \kappa'
    }
  \]
  By the IH on the first premise, \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash T \tpsynth
  \abs{\Pi}{X}{\kappa'}{\kappa}\).
  By the IH on the second premise, \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash T_2
  \tpsynth \kappa''\).
  We also have that the depths of the resulting derivations are no greater than
  the those used to derive the premises, so we conclude using the rule, keeping
  the third premise.

  % type checking
  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash t\tpcheck T
    }{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash t\tpsynth T'
      & T' \cong T
    }
  \]
  By the IH on the first premise, \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash t \tpsynth
  T''\) for some \(T'' \cong T'\).
  By transitivity of convertibility and the second premise, \(T'' \cong T\).
  By transitivity again, \(T'' \cong T_4\) (where \(T_4 \cong T\) is given to
  us).
  We also have that the depths of the resulting derivations are no greater than
  the those used to derive the premises, so we conclude using the rule.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash \absu{\lambda}{x}{t} \tpcheck T
    }{
      T \leadsto_\beta^* \abs{\Pi}{x}{T_5}{T_6}
      & \Gamma_1,x_1:T_1,\Gamma_2,x:T_5\vdash t\tpcheck T_6
    }
  \]
  From the first premise, we have \(T \cong \abs{\Pi}{x}{T_5}{T_6}\), so by
  transitivity of convertibility and assumption, \(T_4 \cong
  \abs{\Pi}{x}{T_5}{T_6}\).
  This means that \(T_4 \leadsto_\beta^* \abs{\Pi}{x}{T_5'}{T_6'}\) for some
  \(T_5' \cong T_5\) and \(T_6' \cong T_6\).
  By one use of the IH on the first premise, we have
  \(\Gamma_1,x_1:T_2,\Gamma_2,x:T_5 \vdash t \tpcheck T_6'\).
  By another use, we have \(\Gamma_1,x_1:T_2,\Gamma_2,x:T_5' \vdash t \tpcheck
  T_6'\) (this is well-founded, because the depth of the typing derivation this
  second use is  applied to is no greater that the depth of the premise).
  We also have that the depths of the resulting derivations are no greater than
  the those used to derive the premises, so we conclude using the rule.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash \absu{\Lambda}{X}{t} \tpcheck T'
    }{
      T' \leadsto_\beta^* \abs{\forall}{X}{\kappa}{T}
      & \Gamma_1,x_1:T_1,\Gamma_2,X:\kappa \vdash t \tpcheck T
    }
  \]
  From the first premise, \(T' \cong \abs{\forall}{X}{\kappa}{T}\).
  From transitivity of convertibility and assumption, \(T_4 \cong
  \abs{\forall}{X}{\kappa}{T}\).
  This means that \(T_4 \leadsto_\beta^* \abs{\forall}{X}{\kappa''}{T''}\) for
  some \(\kappa'' \cong \kappa\) and \(T'' \cong T\).
  By mutual induction with Lemma~\ref{lem:ctxt-conv-class} on the second premise,
  \(\Gamma_1,x_1:T_1,\Gamma_2,X:\kappa'' \vdash t \tpcheck T\), and this
  derivation is no deeper than the second premise.
  By the IH, \(\Gamma_1,x_1:T_1,\Gamma_2,X:\kappa'' \vdash t \tpcheck T''\), and
  this derivation is no deeper than the second premise.
  We conclude using the rule, with the measure preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash [ t , t' ] \tpcheck T_3
    }{
      \begin{array}{cc}
        T_3 \leadsto_\beta^* \abs{\iota}{x}{T}{T'}
        & \Gamma_1,x_1:T_1,\Gamma_2 \vdash t \tpcheck T
        \\ \Gamma_1,x_1:T_1,\Gamma_2 \vdash t' \tpcheck [t/x]^{T}\ T'
        & |t| =_{\beta\eta} |t|
      \end{array}
    }
  \]
  From the first premise, \(T_3 \cong \abs{\iota}{x}{T}{T'}\).
  From transitivity of convertibility and assumption, \(T_4 \cong
  \abs{\iota}{x}{T}{T'}\).
  This means \(T_4 \leadsto \abs{\iota}{x}{T''}{T'''}\) for some \(T'' \cong T\)
  and \(T''' \cong T'\).
  By the IH, \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash t \tpcheck T''\) with depth no
  greater than the second premise (read left to right, then top to bottom).
  By the IH, \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash t' \tpcheck T'''\) with depth no
  greater than the third premise.
  We conclude using the rule, keeping the fourth premise as-is and noting the
  measure is preserved.

  \startcase{.2cm}
  \[
    \infer{
      \Gamma_1,x_1:T_1,\Gamma_2 \vdash \beta\{t'\} \tpcheck T_3
    }{
      T_3 \leadsto_\beta^* \{t_1 \simeq t_2\}
      \quad \textit{FV}(t')\subseteq \textit{dom}(\Gamma_1,x_1:T_1,\Gamma_2)
      \quad |t_1| =_{\beta\eta} |t_2|
    }
  \]
  From the first premise, \(T_3 \cong \{t_1 \simeq t_2\}\).
  From transitivity of the convertibility relation and assumption, \(T_3 \cong
  \{t_1 \simeq t_2\}\).
  This means \(T_4 \leadsto_\beta^* \{t_3 \simeq t_4\}\) for some \(t_3,t_4\)
  such that \(|t_3| =_{\beta\eta} |t_1|\) and \(|t_4| =_{\beta\eta} |t_2|\).
  From the third premise and transitivity of convertibility, we have \(|t_3|
  =_{\beta\eta} |t_4|\).
  From the second premise, we obtain that \(\textit{FV}(t') \subseteq
  \textit{dom}(\Gamma_1,x_1:T_2,\Gamma_2)\).
  We conclude using the rule, noting that the depth remains 1.

  \startcase{.2cm}
  \[
    \infer{\Gamma_1,x_1:T_1,\Gamma_2 \vdash \varphi\ t\ \mhyph\ t'\ \{t''\} \tpcheck T_3}
    {
      \begin{array}{ccc}
        \Gamma_1,x_1:T_1,\Gamma_2 \vdash t\tpsynthleads \{t_1\simeq t_2\}
        &
        & \Gamma_1,x_1:T_1,\Gamma_2\vdash t' \tpcheck T_3
        \\ |t_1| =_{\beta\eta} |t'|
        & |t_2| =_{\beta\eta} |t''|
        & \textit{FV}(t'') \subseteq \textit{dom}(\Gamma_1,x_1:T_1,\Gamma_2)
      \end{array}
    }
  \]
  By the IH on the first premise (first row), \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash
  t \tpsynth T\) (with no greater depth) for some \(T \cong \{t_1 \simeq t_2\}\).
  This means that \(T \leadsto_\beta^* \{t_3 \simeq t_4\}\) for some \(t_3,t_4\)
  such that \(|t_3| =_{\beta\eta} |t_1|\) and \(|t_4| =_{\beta\eta} |t_2|\).
  Using the first and second premise of the second row and transitivity of
  \(\beta\eta\)-equivalence, \(|t_3| =_{\beta\eta} |t'|\) and \(|t_4|
  =_{\beta\eta} |t''|\).
  By the IH on the second premise of the first row, \(\Gamma_1,x_1:T_2,\Gamma_2
  \vdash t' \tpcheck T_4\) (with no greater depth, and where \(T_4\) is the type
  given to us in the lemma and \(T_4 \cong T_3\)).
  Also, we can see from the last premise of the second row that
  \(\textit{FV}(t') \subseteq \textit{dom}(\Gamma_1,x_1:T_2,\Gamma_2)\).
  So, we use the rule to conclude \(\Gamma_1,x_1:T_2,\Gamma_2 \vdash \varphi\ t\
  \mhyph\ t'\ \{t''\} \tpcheck T_4\).
